---
title: "FinalProject"
author: "Nicholas Stout"
date: "November 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
ks_data = read.csv("ks-projects-201801.csv")
```
#Authors
Nicholas Stout, Daniel Tacconi, Issac Nellans

#Proposal
We would like to explore this data to find what effects if a Kickstarter succeeds, looking at amount of time to pledge, the goal amount, and how it is effected by the category.

#Marginal Distributions
First, correction of the date/time variable Launched:
```{r}
ks_data = ks_data %>% separate(launched, into=c("launch_date", "launch_time"), sep = " ")# %>% mutate(launch_date = ymd(launch_date), deadline = ymd(deadline), ID = as.factor(ID))
```
##"category"        
##"main_category"    
##"currency"         
##"deadline"        
##"goal"
##"launch_date"
##"launch_time"     
##"pledged"
##"state"
##"backers"         
##"country"
##"usd.pledged""
##"usd_pledged_real"
##"usd_goal_real"

#Expanding complexity

```{r}

```
