---
title: "FinalProject"
author: "Nicholas Stout"
date: "November 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
ks_data = read.csv("ks-projects-201801.csv")
```
#Authors
Nicholas Stout, Daniel Tacconi, Issac Nellans

#Proposal
We would like to explore this data to find what effects if a Kickstarter succeeds, looking at amount of time to pledge, the goal amount, and how it is effected by the category.

#Marginal Distributions
First, correction of the date/time variable Launched:
```{r}
ks_data = ks_data %>% separate(launched, into=c("launch_date", "launch_time"), sep = " ")# %>% mutate(launch_date = ymd(launch_date), deadline = ymd(deadline), ID = as.factor(ID))
```
##"category"
##"main_category"    
##"currency"         
##deadline       
##goal
##launch_date
##launch_time    
##pledged
##state
##backers        
##country
###Nic Stout
Prediction: I expect most of these to be in the US, by a wide margin
```{r}
summary(ks_data$country)
ks_data %>% ggplot(aes(x=country))+geom_bar()
```

my prediction was correct!

##usd.pledged
###Nic Stout
Prediction: I expect the median to be around 5000, but the mean to be much closer lower than the goal, maybe around 5000 as well. I'm sure the max is a wildly large number that I can't guess of the top of my head.
```{r}
summary(ks_data$usd.pledged)
```

The graph starts starts high and Interestingly, there were some na values, and I was pretty far off. I was right about the max. I did not take into account that the mean would be skewed by the massive high end of this variable.

```{r}
ks_data %>% filter(usd.pledged <= 100000) %>% ggplot(aes(x=usd.pledged))+geom_histogram(binwidth=1000)
```

A vast majority of the amount pledged per kickstarter is less than 1000. This is interesting, as I was expecting it to be more in the few thousands. I am wondering if the amount with none pledged is throwing it off
```{r}
ks_data %>% filter(usd.pledged <= 100000 & usd.pledged > 0) %>% ggplot(aes(x=usd.pledged))+geom_histogram(binwidth=1000)
```

It changed a bit, but still mostly in the < 1000 dollar range. Interesting!

```{r}
ks_data %>% filter(usd.pledged <= 100000 & usd.pledged > 0) %>% ggplot(aes(x=usd.pledged))+geom_density()
mode(ks_data$usd.pledged)
```

This dataset is best describeds as a pareto distribution, as there are far more with a small amount pleged than a large amount. The density graph provided here shows this quite well

##usd_pledged_real
###Nic Stout
Prediction: I expect usd.pledged and usd_pleged_real to be very similar, considering it is the same data with two different conversions.
```{r}
summary(ks_data$usd_pledged_real)
```

That was more different than I was expecting! nearly double on most of the data
```{r}
ks_data %>% filter(usd.pledged <= 100000 & usd.pledged > 0) %>% ggplot(aes(x=usd.pledged))+geom_histogram(binwidth=1000)
```

Very similar histograms so I expect the difference is just due to different conversion rates.

##usd_goal_real
###Nic Stout
prediction: I expect the data to fall between 100 and 100,000, median around 5000, and mean around 25000. I expect to data to be mostly in the < 100000 range
```{r}
summary(ks_data$usd_goal_real)
```
```{r}
ks_data %>% filter(usd_goal_real < 100000) %>% ggplot(aes(x=usd_goal_real))+geom_histogram(binwidth=1000)
```

Ignoring the very large outliers for the sake of visibity, my prediction was fairly close. a distribution with spikes on round numbers (25000, 50000, etc). I was fairly close on the median, but way off on the mean. I did not realize how many values above 1,000,000 there were.

```{}
ks_data %>% filter(usd_goal_real < 100000) %>% ggplot(aes(x=usd_goal_real))+geom_density()
```


#Expanding complexity

```{r}

```
